Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-11-08 chapter_1.Rmd
minor edit. added in headings to discuss later
some minor edits and changes

Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology. Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival. Kelps are able to rapidly adapt their morphological characteristics to reduce drag and avoid dislodgment [@Blanchette1997]. For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and was a combination of *in situ* sampling and transplant of juvenile plants. The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies. Juveniles transplanted into wave exposed sites under went rapid morphological adaption, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaption.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994]. The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments. Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed gentic traits. A study by @?? showed that *species name* was able to rapidly adapt to the local wave environment when transplanted from a low wave exposure site to a high wave exposure site. Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species.**insert example here**. The morphological adaption that macroalgae display are therefore driven by site conditions, therefore individuals must be morphologically flexible to persist in stochastic environments. This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities. For example some algae have fronds and others are articulated corraline, and therefore these species would need to adapt their morphology differently in order to persist. In general, flat strap-like blades are common in areas that are exposed to high wave energy, while at proected sites blade morphology is wide and undulated.

@moss1948 investigated the anatomy, chemical composistion of *Fucus spiralis* at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation in organic nitrogen, mannitol, laminarian and alginic acid concentrations.

@@

>>>>

Temperature and wave exposure have been shown to be important drivers of seaweed distribution, physiological functioning, ability to recover, population dynamics and morphology. Mechanical forces generated by the hydrodynamic environment, in the form of sudden strong ocean currents or storms, between 10- 20 m s-1 with accelerations of 400 m s-2 [@Friedland1995] are the biggest threat to kelp survival. Kelps are able to rapidly adapt their morphological characteristics to reduce drag and avoid dislodgment [@Blanchette1997]. For example a study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Another study by @FowlerWalker2006 tested for differences in morphology of *Ecklonia radiata* between wave-sheltered and wave-exposed sites and through a combination of *in situ* sampling and transplantation of juvenile plants. The results showed that morphology differed between wave-sheltered and wave-exposed sites (thin thallus at sheltered sites and a narrow, thick thallus with a thick stipe at exposed sites), and was consistent with previous studies. Juveniles transplanted into wave exposed sites under went rapid morphological adaptation, whilst the opposite was true for wave-sheltered sites which showed slower morphological adaptation.

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determing distribution, abundance, diversity, composition, growth [@Cousens1982] and productivity [@pedersen2012a] of macroalgae communties. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communties indirectly through the alteration in affect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundry layer around the frond [@lobban1994]. The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgment, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarly in the marine enviroment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macrolagae have been able to persist in often harsh and variable wave environments. Macroalage are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed gentic traits. A study by @Koehletal2008 showed that transplanted *Nereocystis luetkeana* plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delination based on morphology has been inaccurate, and organisms that were once considered two seperate species are actually one species. The morphological adaption that macroalgae display are therefore driven by site conditions, therefore individuals must be morphologically flexible to persist in stochastic environments. This may be achieved through different strategies and are species-specific, which can be directly attributed the high diversity in morphological characters of algal communities. For example some algae have fronds and others are articulated corraline, and therefore these species would need to adapt their morphology differently in order to persist. In general, flat strap-like blades are common in areas that are exposed to high wave energy, while at proected sites blade morphology is wide and undulated.

@@

# Coastal flow

# Kelp-rafting
